Forecasting the 2024 U.S. Presidential Election

### 2024-11-02 paper/paper.qmd

library(tidyverse)
library(ggplot2)
library(janitor)
library(modelsummary)
library(here)
library(dplyr)
library(knitr)
library(car)

@@

```{r}
#| echo: false
#| message: false
#| warning: false
# Load necessary library
library(ggplot2)

@@

```{r}
#| echo: false
#| message: false
#| warning: false
# Load necessary libraries
library(reshape2)
library(ggplot2)

@@

# Load necessary libraries
library(ggplot2)
library(reshape2)
library(dplyr)
library(patchwork)

@@

2. **Aggregation of Predicted Support by State**: For each candidate, we computed the average predicted polling percentage within each state. This aggregation provides a concise state-level summary of anticipated support, facilitating direct comparisons between the candidates in each state (see Figure @fig-predicted-vote-pct).

@@

7. **Interpretation of Results**: The analysis indicated that Kamala Harris is projected to receive a total of **306 electoral votes**, while Donald Trump is projected to receive **232 electoral votes**. Harris surpasses the 270-vote threshold required for victory, suggesting a strong position in the Electoral College based on our predictions. This outcome implies that Harris is likely to win the election, given her substantial lead in projected electoral votes.

@@

>>>>

library(tidyverse)
library(ggplot2)
library(janitor)
library(modelsummary)
library(here)
library(dplyr)
library(knitr)
library(car)
library(maps)
library(stringr)
library(reshape2)
library(patchwork)

@@

```{r}
#| echo: false
#| message: false
#| warning: false
# Load necessary library

@@

```{r}
#| echo: false
#| message: false
#| warning: false
# Load necessary libraries

@@

@@

2. **Aggregation of Predicted Support by State**: For each candidate, we computed the average predicted polling percentage within each state. This aggregation provides a concise state-level summary of anticipated support, facilitating direct comparisons between the candidates in each state (see @fig-predicted-vote-pct).

@@

7. **Interpretation of Results**: The analysis indicated that Kamala Harris is projected to receive a total of **306 electoral votes**, while Donald Trump is projected to receive **232 electoral votes**. Harris surpasses the 270-vote threshold required for victory, suggesting a strong position in the Electoral College based on our predictions. This outcome implies that Harris is likely to win the election, given her substantial lead in projected electoral votes (see @fig-map-state-winner).

@@

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-map-state-winner
#| fig.cap: "Electoral College Predition Map"

# Get US state map data
us_states <- map_data("state")

# Create a mapping of state names
state_mapping <- data.frame(
  state = c(
    "Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado", "Connecticut", "Delaware",
    "District of Columbia", "Florida", "Georgia", "Hawaii", "Idaho", "Illinois", "Indiana", "Iowa",
    "Kansas", "Kentucky", "Louisiana", "Maine At-Large", "Maine CD-1", "Maine CD-2", "Maryland",
    "Massachusetts", "Michigan", "Minnesota", "Mississippi", "Missouri", "Montana",
    "Nebraska At-Large", "Nebraska CD-1", "Nebraska CD-2", "Nebraska CD-3",
    "Nevada", "New Hampshire", "New Jersey", "New Mexico", "New York", "North Carolina",
    "North Dakota", "Ohio", "Oklahoma", "Oregon", "Pennsylvania", "Rhode Island",
    "South Carolina", "South Dakota", "Tennessee", "Texas", "Utah", "Vermont",
    "Virginia", "Washington", "West Virginia", "Wisconsin", "Wyoming"
  ),
  region = tolower(c(
    "alabama", "alaska", "arizona", "arkansas", "california", "colorado", "connecticut", "delaware",
    "district of columbia", "florida", "georgia", "hawaii", "idaho", "illinois", "indiana", "iowa",
    "kansas", "kentucky", "louisiana", "maine", "maine", "maine", "maryland",
    "massachusetts", "michigan", "minnesota", "mississippi", "missouri", "montana",
    "nebraska", "nebraska", "nebraska", "nebraska",
    "nevada", "new hampshire", "new jersey", "new mexico", "new york", "north carolina",
    "north dakota", "ohio", "oklahoma", "oregon", "pennsylvania", "rhode island",
    "south carolina", "south dakota", "tennessee", "texas", "utah", "vermont",
    "virginia", "washington", "west virginia", "wisconsin", "wyoming"
  ))
)

# Simplify prediction data for mapping
map_data <- prediction_comparison %>%
  select(state, winner) %>%
  left_join(state_mapping, by = "state")

# Merge with map data
map_data_final <- us_states %>%
  left_join(map_data, by = "region")

# Create the basic map without projection
ggplot(data = map_data_final) +
  geom_polygon(aes(x = long, y = lat, group = group, fill = winner), 
               color = "white", size = 0.2) +
  coord_fixed(1.3) +  # This replaces coord_map() with a simpler aspect ratio adjustment
  scale_fill_manual(values = c("Harris" = "#87CEFA", "Trump" = "#F08080"),
                    name = "Predicted Winner") +
  theme_minimal() +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    plot.title = element_text(hjust = 0.5, size = 16),
    legend.position = "bottom"
  )
```
